---
title: "Hello, `dygraphs`"
author: "Jeff Newmiller"
date: "April 2, 2019"
output: slidy_presentation
---

```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.height = 3)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages( library(xts) )
library(dygraphs)
```

## `dygraphs` Package

- Generates interactive time-series plots for embedding in HTML
- Works in "static" html (self-contained file) as well as "dynamic" html (Shiny)
- Uses the [dygraphs](https://rstudio.github.io/dygraphs/) JavaScript library, embeds data into the html
- MIT licensed

```{r,eval=FALSE}
install.packages("dygraphs",dependencies = TRUE)
library(xts)
library(dygraphs)
```


## Basic Plot

`nhtemp` is a data set in the `datasets` package that comes with R.

```{r}
dygraph(nhtemp, main = "New Haven Temperatures")
```

## How does it work?

The `dygraph` function returns a "dygraphs" object, which is a kind of "htmlwidget".

When you print a `dygraph` object at the RStudio console it recognizes that it is an "htmlwidget" and displays it in the Viewer pane. (Convenient!)

When you print it from within a `knitr` code chunk, `knitr` recognizes that it is an "htmlwidget" and renders the JavaScript and data into the output file. Note that this use case does not **require** RStudio, which can be convenient if you want to generate graphs on a headless server.

When a web browser runs the script (later), it becomes interactive.

## Grouped Graphs

```{r,echo=TRUE,fig.height=1}
dygraph(ldeaths, main = "All", group = "lung-deaths")
dygraph(mdeaths, main = "Male", group = "lung-deaths")
dygraph(fdeaths, main = "Female", group = "lung-deaths")
```

## Alternate Zoom

```{r}
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector()
```


## Candlestick Display

```{r}
library(xts)
data(sample_matrix)
m <- tail(sample_matrix, n = 32)
dygraph(m) %>%
  dyCandlestick()
```

## Inputs

The `dygraph` function expects an `xts` object or a data frame with a `numeric` first column.

```{r}
Sys.setenv( TZ="GMT" )
dta <- read.csv( "../data/MAC000002clean.csv"
               , as.is = TRUE
               , colClasses = c( NA, "POSIXct", "numeric" )
               )
Sys.setenv( TZ="Europe/London" )
# as.matrix keeps column label from data frame
dtaxts <- xts( as.matrix( dta[ , "KWH", drop=FALSE ] )
             , order.by = dta$DateTime
             )
dygraph( dtaxts
       , main = "Sample Home Electricity Consumption"
       , ylab = "Power (kWh)"
       ) %>% 
  dyRangeSelector()
```

## Conclusion

- Ideal for timeseries, not a general purpose solution
- https://rstudio.github.io/dygraphs/
- `install.packages("dygraphs")`
